Analysis of epistatic inflation under large single additive effects, with a focus on the replication rate

### 2020-01-21 null_replication/docs/analyse_results.rmd

```{r}
library(tidyverse)
load("../data/rs67903230.rdata")
res$polygenic <- res$i %% 2 == 0
```

@@

```{r}
hist(res$varexp)
```

@@

```{r}
plot(lambda_disc ~ lambda_repl, res)
```

@@

```{r}
ggplot(res, aes(x=varexp, y=lambda_disc)) +
geom_point(aes(colour=polygenic)) +
geom_smooth(aes(colour=polygenic))
```

@@

```{r}
ggplot(res, aes(x=nsig_disc_bonf)) +
geom_histogram(aes(fill=polygenic), position="dodge")
```

@@

```{r}
ggplot(res, aes(x=varexp, y=nsig_disc_bonf)) +
geom_point()
```

@@

@@

```{r}
res$lambda_disc_bin <- Hmisc::cut2(res$lambda_disc, g=30)
group_by(res, lambda_disc_bin) %>% summarise(n=mean(nsig_disc_bonf, na.rm=T)) %>% 
ggplot(aes(x=lambda_disc_bin, y=n)) +
geom_point() +
theme(axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5)) +
labs(x="Simulation's lambda", y="Mean number of discovery hits")
```

@@

```{r}
res$bonf_bonf_rr <- res$cond_bonf_bonf / res$nsig_disc_bonf
res$bonf_fdr_rr <- res$cond_bonf_fdr / res$nsig_disc_bonf
res$bonf_bonf2_rr <- res$cond_bonf_bonf2 / res$nsig_disc_bonf

repl_rate <- bind_rows(
	group_by(res, lambda_disc_bin) %>% summarise(m=mean(bonf_bonf_rr, na.rm=T)) %>% mutate(threshold="Bonferroni"),
	group_by(res, lambda_disc_bin) %>% summarise(m=mean(bonf_fdr_rr, na.rm=T)) %>% mutate(threshold="FDR"),
	group_by(res, lambda_disc_bin) %>% summarise(m=mean(bonf_bonf2_rr, na.rm=T)) %>% mutate(threshold="Experiment")
)
repl_rate %>% ggplot(aes(x=lambda_disc_bin, y=m, group=threshold)) +
geom_line(aes(colour=threshold)) +
scale_colour_brewer(type="qual") +
theme(axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5)) +
labs(x="Simulation's lambda", y="Replication rate")
```

>>>>

```{r}
l <- list()
# load("../data/rs67903230.rdata")
# res$polygenic <- res$sim %% 2 == 0
# res$median_p_disc <- pchisq(res$lambda_disc * qchisq(0.5, 1), 1, low=FALSE)
# res$median_p_repl <- pchisq(res$lambda_repl * qchisq(0.5, 1), 1, low=FALSE)
# res$lambda_disc <- pnorm(1 - res$median_p_disc / 2)^2 / qchisq(0.5, 1)
# res$lambda_repl <- pnorm(1 - res$median_p_repl / 2)^2 / qchisq(0.5, 1)
# res$gene <- "MBNL1"
# res$i <- paste(res$gene, res$sentinel, res$cis, sep="_")
# res$scenario <- 1

load("../data/aggregate1.rdata")
res$polygenic <- res$sim %% 2 == 0
res$gene <- sapply(res$i, function(x) strsplit(x, "_")[[1]][1])
res$scenario <- 1
l[[1]] <- res

load("../data/aggregate.rdata")
res$polygenic <- res$sim %% 2 == 0
res$gene <- sapply(res$i, function(x) strsplit(x, "_")[[1]][1])
res$scenario <- 2
l[[2]] <- res

load("../data/aggregate2.rdata")
res$polygenic <- res$sim %% 2 == 0
res$gene <- sapply(res$i, function(x) strsplit(x, "_")[[1]][1])
res$scenario <- 3
l[[3]] <- res

res <- bind_rows(l)

```

@@

```{r}
ggplot(res, aes(x=varexp)) +
geom_histogram() +
facet_grid(scenario ~ .)
```

@@

```{r}
ggplot(res, aes(y=lambda_repl, x=lambda_disc)) +
geom_point() +
labs(x = "Genomic inflation (discovery)", y = "Genomic inflation (replication)") +
facet_grid(scenario ~ .)
```

@@

```{r}
ggplot(res, aes(x=varexp, y=lambda_disc)) +
geom_point(size=0.03, alpha=0.1) +
labs(x="Additive variance explained", y="Genomic inflation") +
facet_grid(scenario ~ .) +
theme_bw()
```

@@

```{r}
ggplot(res, aes(x=nsig_disc_bonf)) +
geom_histogram(aes(fill=polygenic), position="dodge") +
facet_grid(scenario ~ .)
```

@@

```{r}
res$varexp_bin <- Hmisc::cut2(res$varexp, g=30)
ggplot(res, aes(x=varexp_bin, y=nsig_disc_bonf)) +
geom_boxplot() +
theme(axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5)) +
labs(x="Additive variance explained", y="Mean number of discovery hits") +
facet_grid(scenario ~ .)
```

@@

```{r}
ggplot(res, aes(x=fstat_cor)) +
geom_histogram(fill="white", colour="black") +
labs(x="Correlation of F statistics between discovery and replication") +
facet_grid(scenario ~ ., scale="free_y")
```

@@

```{r}
res$lambda_disc_bin <- Hmisc::cut2(res$lambda_disc, g=30)
res$varexp_bin <- Hmisc::cut2(res$varexp, g=30)
res %>% 
ggplot(aes(x=lambda_disc_bin, y=nsig_disc_bonf)) +
geom_boxplot() +
theme(axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5)) +
labs(x="Genomic inflation of simulation", y="Mean number of discovery hits") +
facet_grid(scenario ~ .)

res %>% 
ggplot(aes(x=varexp_bin, y=nsig_disc_bonf)) +
geom_boxplot() +
theme(axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5)) +
labs(x="Variance explained by cis causal variant", y="Mean number of discovery hits") +
facet_grid(scenario ~ .)
```

@@

```{r, fig.width=12}
res$bonf_bonf_rr <- res$cond_bonf_bonf / res$nsig_disc_bonf
res$bonf_fdr_rr <- res$cond_bonf_fdr / res$nsig_disc_bonf
res$bonf_bonf2_rr <- res$cond_bonf_bonf2 / res$nsig_disc_bonf

repl_rate <- bind_rows(
	group_by(res, varexp_bin, scenario) %>% summarise(m=mean(bonf_bonf_rr, na.rm=T)) %>% mutate(threshold="Bonferroni"),
	group_by(res, varexp_bin, scenario) %>% summarise(m=mean(bonf_fdr_rr, na.rm=T)) %>% mutate(threshold="FDR"),
	group_by(res, varexp_bin, scenario) %>% summarise(m=mean(bonf_bonf2_rr, na.rm=T)) %>% mutate(threshold="Experiment")
)
repl_rate %>% ggplot(aes(x=varexp_bin, y=m, group=threshold)) +
geom_line(aes(colour=threshold)) +
scale_colour_brewer(type="qual") +
theme(axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5)) +
labs(x="Variance explained by cis additive variant", y="Replication rate") +
facet_grid(. ~ scenario) +
ylim(c(0,0.3))
```
